SPENCER THESIS - WORK IN PROGRESS

### 2016-05-26 doc/thesis_swe.tex
Thesis update.

Intro\\

Place in context of recent work\\

The goals of this thesis are two answer the two following questions: How accurately can we reconstruct line-of-sight dark matter mass structures using a halo model? From our model outputs, how feasible is an inference of the stellar mass – halo mass relation parameters?\\

Outline thesis

% OLD: In a universe teeming with galaxies and light, it came as a shock when 20\textsuperscript{th} century astronomers discovered that most of the mass in the universe is in fact dark; the `normal' matter made of atoms that we interact with in everyday life, called baryonic matter, accounts for less than 20\% of the mass in the observable universe. The remaining mass takes the form of an exotic dark matter that does not absorb or emit light, rendering it invisible to our telescopes. While this claim sounds bizarre, there has been an abundance of indirect evidence in recent decades for the existence of dark matter including the flattening of galaxy rotation curves \cite{rotation_curve}, velocity dispersions of galaxies \cite{zwicky}, and acoustic peaks in the cosmic microwave background \cite{planck_2015}.

% One of the most successful probes of dark matter has been gravitational lensing. The path of light from distant `background' galaxies is bent when traveling through regions of space containing large amounts of `foreground' mass. Light from different origins in a source galaxy is subject to different bending which results in a distortion of the galaxy image. As the foreground mass is known to be largely dark matter, gravitational lensing supplies a direct constraint on the mass of dark matter in that region.

% While the effects of gravitational lensing can be dramatic, the shape of most galaxies is only distorted by a few percent and must be detected statistically as the intrinsic shape is not known. If a model for the distribution of dark matter in a region of foreground mass can accurately predict the statistical signal of this `weak' lensing of background sources, then the model can be used on galaxies in existing sky survey data to extrapolate the amount of dark matter in the region and construct large-scale maps of the dark matter in the universe.

% In this paper, I attempt to do the former by applying a simple dark matter halo model to reconstruct mass along lines of sight in the Millennium Simulation and predict the weakly lensed ellipticities of generated background sources. A brief introduction into the theory of galaxy ellipticities, the effects of strong and weak gravitational lensing, and dark matter halos is discussed in Section II. The implementation of a halo model on the Millennium Simulation is described in Section III, and the results of the model on 3,240 galaxies on a field size of 324 arcmin$^2$, along with a comparison of the predicted lensed ellipticities to ray-traced ellipticities, is given in Section IV.  Section V discusses limitations of the used framework as well as potential physical and computational improvements that can be made for upcoming work before concluding remarks in Section VI.

@@

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% FIGURE: Millennium Simulation Distribution plots (cosmic web, galaxy clusters with/without dark matter)
% \begin{figure*}
%     \centering
%     \begin{subfigure}[H]{0.4\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/thesis/ms_galaxies.jpg}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{CAPTION}
%         \label{ms_galaxies}
%     \end{subfigure}
%     ~
%     %% Einstein Ring Figure
%     \begin{subfigure}[H]{0.4\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/thesis/ms_halos.jpg}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{CAPTION}
%         \label{ms_halos}
%     \end{subfigure}
%     \caption{}
% \end{figure*}

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Ellipticity Orientation Figure
% \begin{figure*}
%     \centering
%     \begin{subfigure}[H]{0.415\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/ellipse_orientations.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{The shape of a galaxy image for various ellipticity components $\varepsilon_1$ and $\varepsilon_2$ on the $x$ and $y$-axes respectively. Taken from \cite{schneider}.}
%         \label{ellipses}
%     \end{subfigure}
%     ~
%     %% Einstein Ring Figure
%     \begin{subfigure}[H]{0.425\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/einstein_ring.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{An image of the luminous red galaxy LRG 3-757 along with a strongly lensed background galaxy, called the `Cosmic Horseshoe' \cite{einstein_ring}.}
%         \label{einstein_ring}
%     \end{subfigure}
%     \caption{}
% \end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Pangloss Cartoon
% \begin{figure*}
%     \centering
%     \includegraphics[width=0.85\textwidth]{figs-swe/pangloss_cartoon.png}
%     \captionsetup{justification=raggedright,singlelinecheck=false}
%     \caption{A cartoon model of how the \texttt{Pangloss} framework uses a dark matter halo model to reconstruct the foreground mass along the line of sight of a background source galaxy. A `lightcone' centered at a background source is constructed with radius $R$ and populated with all foreground galaxies contained in this volume. Each foreground object has an attached dark matter halo and physical distance to the line of sight (dashed line). The shear and convergence along the line of sight contributed by each halo is calculated using \cite{lensing_calc}, and the predicted ellipticity is determined using the sum of these contributions and Equation \eqref{lensing}.}
%     \label{pangloss_cartoon}
% \end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

where $M_h$ is the halo mass and $R$ is the distance from the halo to the line of sight. The relevance distribution for one background catalog realization with \textcolor{red}{???} foreground halos in a lightcone of radius \textcolor{red}{??} arcminutes is shown in Figure ??. As expected, there are very few massive halos along the line of sight and therefore only a small percentage of foreground halos should have a high relative contribution to the predicted lensing quantities. This encouraged the intended analysis on lightcones with all halos and those with only a small subset of the most relevant halos to compare the results and see if the computational time saved by using only the most relevant halos was worth the loss in prediction accuracy.

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Relevance Distribution of all Halos
% \begin{figure}
%     \centering
%     \includegraphics[width=0.25\textwidth]{figs-swe/thesis/relevance_distribution.png}
%     \captionsetup{justification=raggedright,singlelinecheck=false}
%     \caption{CAPTION}
%     \label{rel_dist}
% \end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Smooth-Component Correction Diagram
% \begin{figure}
%     \centering
%     \includegraphics[width=0.25\textwidth]{figs-swe/thesis/smooth_correction.png}
%     \captionsetup{justification=raggedright,singlelinecheck=false}
%     \caption{CAPTION}
%     \label{scc}
% \end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

As discussed in Section \ref{scc}, the halo model does not account for additional mass structures such as filaments. To apply the smooth-component correction to \texttt{Pangloss}, the mean mass density $\rho(z)$ and halo density $\rho_h(z)$ had to be calculated at each of the 100 redshift bins. For $\rho_h(z)$, this was done by summing all halo masses $M_{h,i}$ in a given redshift bin and integrating Equation \eqref{pvolume_element} from ${z-\Delta z/2}$ to ${z+\Delta z/2}$, except for the boundary bins. Once $\rho_h(z)$ had been calculated for each redshift slice, the projected surface mass density was found by integrating over the redshift interval,
\begin{equation}\label{surface_density} 
\sum(z)=\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \rho_s(z)\,dz=\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \rho(z)-\rho_h(z)\,dz,
\end{equation}

\noindent and the convergence contribution of the mass sheet was calculated using Equation \eqref{convergence} with ${z_l=z}$ and ${z_s=1.3857}$. The corrected total convergence for a given background source is simply the sum of the halo contributions and the asfd. There is no additional shear produced by a uniform sheet of mass (citation!), but the corrected reduced shear is now calculated using the corrected total $\kappa$ in Equation \eqref{reduced_shear}. A diagram of this process is shown in Figure \ref{fig:scc}.

The volume and projected surface mass densities of each redshift slice are shown in Figures \ref{fig:volume_densities} and \ref{fig:surface_densities} respectively. (Describe plots)

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Surface and Volume densities for smooth-component correction redshift slices
% \begin{figure*}
%     \centering
%     \begin{subfigure}[H]{0.415\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/thesis/scc_vol_den.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{CAPTION}
%         \label{fig:volume_densities}
%     \end{subfigure}
%     ~
%     %% Einstein Ring Figure
%     \begin{subfigure}[H]{0.425\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/thesis/scc_surf_deng.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{CAPTION}
%         \label{fig:surface_densities}
%     \end{subfigure}
%     \caption{}
% \end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

(Predict LOS kappa and shear measurements )

Using the physical separation distance and halo mass, the shear and convergence contribution of a single foreground halo is calculated using methods described in Wright and Brainerd \cite{lensing_calc}. The total convergence and shear at the center of the lightcone is simply the sum of the convergence and shear contributions of each halo contained in the lightcone. The \texttt{Pangloss} predicted lensed source ellipticity is then calculated using Equation \eqref{lensing}.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Pangloss Lightcone Data
% \begin{figure}
%     \centering
%     \includegraphics[width=0.25\textwidth]{figs-swe/thesis/lightcone_data.png}
%     \captionsetup{justification=raggedright,singlelinecheck=false}
%     \caption{CAPTION}
%     \label{rel_dist}
% \end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

Figure (??) displays different views an example lightcone where the foreground halos are plotted as circles whose size are proportional to their galaxy brightness (top left), angular size (top center), convergence contributions (top right), and halo mass (bottom). Figure (??) shows a few examples of mass structures in the foreground catalog along with the ray-traced convergence and shear maps from (Hilbert cite).

Figure (??) shows a

Instead of comparing the ray-traced and \texttt{Pangloss} predicted ellipticities for individual galaxies, the lensing is characterized globally with correlation functions. As described in section (??), the ellipticity-ellipticity correlation function measures how correlated the ellipticities of pairs of galaxies are as a function of separation distance, while the galaxy-mass correlation function measures the correlation of lensed ellipticities around foreground halos as a function of separation distance. Both correlation functions are used in this work to estimate how well the \texttt{Pangloss} framework models weak lensing by dark matter structures and are computed using the publicly available \texttt{TreeCorr} module written by Mike Jarvis\footnote{https://github.com/rmjarvis/TreeCorr}. Note that the ellipticity-ellipticity correlation function definition used in \texttt{TreeCorr} is slightly different than that used in most of the literature; for a derivation of the connection between Jarvis's definition and the more common Schneider definition \cite{schneider}, see Appendix (??).

% For readers that are unfamiliar with correlation functions in the context of cosmology or want a visual aid, see the appendix.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Ellipticity-Ellipticity Correlation
% \begin{figure}
%     \centering
%     \begin{subfigure}{0.475\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/gg_corr.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{The ellipticity-ellipticity correlation function for 3,240 sources at separation distances from 0.1 to 2 arcminutes and a lightcone radius of $R=2$ arcminutes. While the correlation of the intrinsic ellipticities before lensing (blue) shows no correlation, the correlation of the ray-traced ellipticities (green) and halo model predicted ellipticities (purple) both show significant positive correlation on scales up to 1 arcminute. The halo model predicts a small negative correlation above this scale, while the ray-tracing shows positive correlation on all scales.}
%         %The correlation of the intrinsic ellipticities before lensing is plotted in blue, and shows no correlation. The correlation of the true ray-traced ellipticities is plotted in green and shows significant positive correlation on all scales. The halo predicted correlation, plotted in purple, shows significant positive correlation until a separation distance of 0.8 arcminutes.
%         \label{gg_corr}
%     \end{subfigure}
%     ~~
%     \begin{subfigure}{0.475\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/gg_corr_series.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{The ellipticity-ellipticity correlation functions for a series of halo model frameworks with increasing lightcone radii from $R=1$ to $R=8$ arcminutes, all plotted in purple with various line styles. These are compared to the same ray-traced correlation from Figure \ref{gg_corr}.}
%         \label{gg_corr_series}
%     \end{subfigure}
%     \caption{}
% \end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Galaxy-mass correlation function
% \begin{figure}
%     \centering
%     \includegraphics[width=0.475\textwidth]{figs-swe/ng_corr.png}
%     \captionsetup{justification=raggedright,singlelinecheck=false}
%     \caption{The galaxy-mass correlation function of 3,240 sources at separation distances from 0.1 to 2 arcminutes and a lightcone radius of $R=2$ arcminutes. There is no significant correlation of the intrinsic ellipticities (blue) as expected, and both the ray-traced correlation (green) and halo model correlation (purple) are positive on all scales. The green and purple curves are in agreement from 0.2 to 0.7 arcminutes, but have significant discrepancies on scales above and below this range.}
%     \label{ng_corr}
% \end{figure}
%
% ...Same progression corr function as gg??
%
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% 3 panel plot of binned Hilbert map, Pangloss map, and difference
% \begin{figure}
%     \centering
%     \includegraphics[width=0.475\textwidth]{figs-swe/thesis/kappamaps.png}
%     \captionsetup{justification=raggedright,singlelinecheck=false}
%     \caption{CAPTION}
%     \label{corr_fraction}
% \end{figure}
%
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Mean lightcone CPU time
% \begin{figure}
%     \centering
%     \includegraphics[width=0.475\textwidth]{figs-swe/profile_bar.png}
%     \captionsetup{justification=raggedright,singlelinecheck=false}
%     \caption{The mean CPU runtime per lightcone for various runs with different (and cumulative) code optimizations described in the text. The cumulative changes account for a 70\% decrease in the mean CPU time and a halving of the variation. Currently, the prediction can be made for a single lightcone in\\$24\pm5$ ms.}
%     \label{cpu_plot}
% \end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

% OLD: While the current \texttt{Pangloss} framework can handle the lensing predictions of 3,240 sources in a 324 arcmin$^2$ field comfortably on a single processor, the goal is to scale up the framework to make predictions for background sources of the same number density in a 100 deg$^2$ field; this will require 3.6 million lightcones, each containing up to 1,000 foreground halos. Luckily, the prediction is trivially parallelizable as the shear and convergence calculation of each lightcone is independent of all other lightcones. This makes GPUs an attractive candidate for future work, as it would only take 360 GPUs with 10,000 threads each to carry out the prediction. Additionally, the 50 MB of RAM per lightcone is sufficiently small to fit 10,000 lightcones on a GPU.

@@

% OLD: I would like to thank the Department of Energy for funding this research and giving me this wonderful opportunity, as well as SLAC National Laboratory for hosting me. This work was greatly aided by numerous conversations with Risa Wechsler from Stanford University and Matt Becker from the Kavli Institute for Particle Astrophysics and Cosmology (KIPAC). Finally, I am incredibly grateful to Phil Marshall for his unending support, patience, and guidance throughout this research.
% This work was supported in part by the U.S. Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists (WDTS) under the Science Undergraduate Laboratory Internships Program (SULI).

@@

(Introduce issue a bit)\\

>>>>

In a universe teeming with galaxies and light, it came as a shock when 20\textsuperscript{th} century astronomers discovered that most of the mass in the universe is in fact dark; the `normal' matter made of atoms that we interact with in everyday life, called baryonic matter, accounts for less than 20\% of the mass in the observable universe. The remaining mass takes the form of an exotic dark matter that does not absorb or emit light, rendering it invisible to our telescopes. While this claim sounds bizarre, there has been an abundance of indirect evidence in recent decades for the existence of dark matter including the flattening of galaxy rotation curves \cite{rotation_curve}, velocity dispersions of galaxies \cite{zwicky}, and acoustic peaks in the cosmic microwave background \cite{planck_2015}.

One of the most successful probes of dark matter has been gravitational lensing. The path of light from distant `background' galaxies is bent when traveling through regions of space containing large amounts of `foreground' mass. Light from different origins in a source galaxy is subject to different bending which results in a distortion of the galaxy image. As the foreground mass is known to be largely dark matter, gravitational lensing supplies a direct constraint on the mass of dark matter in that region.

While the effects of gravitational lensing can be dramatic, the shape of most galaxies is only distorted by a few percent and must be detected statistically as the intrinsic shape is not known. If a model for the distribution of dark matter in a region of foreground mass can accurately predict the statistical signal of this `weak' lensing of background sources, then the model can be used on galaxies in existing sky survey data to extrapolate the amount of dark matter in the region and construct large-scale maps of the dark matter in the universe.\\

(Connect to previous work done by Collett et al., Hilbert et al., McCully et al., Pache et al., etc.)\\

%By applying a simple dark matter halo model to reconstruct lensing mass along lines of sight in the Millennium Simulation to predict the weakly lensed ellipticities of generated background sources, in this thesis I attempt to answer the following questions:

In this thesis, I apply a simple dark matter halo model to reconstruct lensing mass along lines of sight in the Millennium Simulation to predict the weakly lensed ellipticities of generated background sources. By doing so, I attempt to answer the following questions: Under the best possible circumstances of having noiseless redshifts and halo masses, how accurately can I reconstruct line-of-sight dark matter mass structures using a simple halo model? What fraction of dark matter mass is in halos? How feasible is an inference of the Millennium Simulation's cosmological hyper-parameters, such as the stellar mass-halo mass relation, using all available halo data? How long does such an inference take?

A brief introduction to the halo model, the effects of strong and weak gravitational lensing, representing galaxy ellipticities, and cosmological correlation functions is discussed in Section II. The implementation of the \texttt{Pangloss} framework to make line of sight lensing predictions in the Millennium Simulation is described in Section III, and the results of the framework on ?? galaxies on a field size of ?? arcmin$^2$, along with a comparison of the predicted lensed ellipticities to ray-traced ellipticities, is given in Section IV. Section V explores the feasibility of using \texttt{Pangloss} to make parameter inferences using all available halo data either by likelihood estimation or approximate Bayesian computation. Section VI discusses limitations of the used framework as well as potential physical and computational improvements that can be made for upcoming work before concluding remarks in Section VII.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% FIGURE: Millennium Simulation Distribution plots (cosmic web, galaxy clusters with/without dark matter)
\begin{figure*}
    \centering
    \begin{subfigure}[H]{0.375\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ms_galaxies.jpg}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{ms_galaxies}
    \end{subfigure}
    ~
    %% Einstein Ring Figure
    \begin{subfigure}[H]{0.375\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ms_halos.jpg}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{ms_halos}
    \end{subfigure}
    \caption{}
\end{figure*}

@@

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Ellipticity Orientation Figure
\begin{figure*}
    \centering
    \begin{subfigure}[H]{0.415\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/ellipse_orientations.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{The shape of a galaxy image for various ellipticity components $\varepsilon_1$ and $\varepsilon_2$ on the $x$ and $y$-axes respectively. Taken from \cite{schneider}.}
        \label{ellipses}
    \end{subfigure}
    ~
    %% Einstein Ring Figure
    \begin{subfigure}[H]{0.425\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/einstein_ring.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{An image of the luminous red galaxy LRG 3-757 along with a strongly lensed background galaxy, called the `Cosmic Horseshoe' \cite{einstein_ring}.}
        \label{einstein_ring}
    \end{subfigure}
    \caption{}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Pangloss Cartoon
\begin{figure*}
    \centering
    \includegraphics[width=0.85\textwidth]{figs-swe/thesis/pangloss_cartoon.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{A cartoon model of how the \texttt{Pangloss} framework uses a dark matter halo model to reconstruct the foreground mass along the line of sight of a background source galaxy. A `lightcone' centered at a background source is constructed with radius $R$ and populated with all foreground galaxies contained in this volume. Each foreground object has an attached dark matter halo and physical distance to the line of sight (dashed line). The shear and convergence along the line of sight contributed by each halo is calculated using \cite{lensing_calc}, and the predicted ellipticity is determined using the sum of these contributions and Equation \eqref{lensing}.}
    \label{pangloss_cartoon}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

\noindent where $M_h$ is the halo mass and $R$ is the distance from the halo to the line of sight. Numerical values are assigned by comparing a halo's individual $M_h$ and $R$ to a threshold $M_T$ and $R_T$:
\begin{equation}\label{relevant_thresh}
Rel(M_h,R)=\frac{(M_h/M_T)}{(R/R_T)^3}.
\end{equation}

In this work, the threshold values used are $M_T=10^{12}M_\odot$ and $R_T=10$ kpc. The relevance distribution for one background catalog realization with \textcolor{red}{???} foreground halos in a lightcone of radius \textcolor{red}{??} arcminutes is shown in Figure ??. As expected, there are very few massive halos along the line of sight and therefore only a small percentage of foreground halos should have a high relative contribution to the predicted lensing quantities. This encouraged the intended analysis on lightcones with all halos and those with only a small subset of the most relevant halos to compare the results and see if the computational time saved by using only the most relevant halos was worth the loss in prediction accuracy.

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Relevance Distribution of all Halos
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/thesis/relevance_distribution.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:rel_dist}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Smooth-Component Correction Diagram
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{figs-swe/thesis/smooth_correction.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:scc}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

As discussed in Section \ref{scc}, the halo model does not account for additional mass structures such as filaments. To apply the smooth-component correction to \texttt{Pangloss}, the mean mass density $\rho(z)$ and halo density $\rho_h(z)$ had to be calculated at each of the 100 redshift bins. For $\rho_h(z)$, this was done by summing all halo masses $M_{h,i}$ in a given redshift bin and integrating Equation \eqref{pvolume_element} from ${z-\Delta z/2}$ to ${z+\Delta z/2}$, except for the boundary bins. Once $\rho_h(z)$ had been calculated for each redshift slice, the projected surface mass density was found by integrating over the redshift interval,
\begin{align}\label{surface_density} 
\sum(z)&=\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \rho_s(z)\,dz\nonumber\\
&=\int_{z-\frac{\Delta z}{2}}^{z+\frac{\Delta z}{2}} \rho(z)-\rho_h(z)\,dz,
\end{align}

\noindent and the convergence contribution of the mass sheet was calculated using Equation \eqref{convergence} with ${z_l=z}$ and ${z_s=1.3857}$. The corrected total convergence for a given background source is simply the sum of the halo contributions and the smooth-component contributions. There is no additional shear produced by a uniform sheet of mass (citation!), but the corrected reduced shear is now calculated using the corrected total $\kappa$ in Equation \eqref{reduced_shear}. A diagram of this process is shown in Figure \ref{fig:scc}.

The volume and projected surface mass densities of each redshift slice are shown in Figures \ref{fig:volume_densities} and \ref{fig:surface_densities} respectively. (Describe plots) (reduced shears before and after correction)

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Surface and Volume densities for smooth-component correction redshift slices
\begin{figure*}
    \centering
    % Volume
    \begin{subfigure}[H]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/vol_densities.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{fig:volume_densities}
    \end{subfigure}
    ~
    % Surface
    \begin{subfigure}[H]{0.49\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/thesis/surf_densities.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        \label{fig:surface_densities}
    \end{subfigure}
    \caption{}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Kappa Contributions of S.C.C.
\begin{figure}
    \centering
    \includegraphics[width=0.425\textwidth]{figs-swe/thesis/kappa_distribution.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:kappa_dist}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

Using the physical separation distance and halo mass, the shear and convergence contribution of a single foreground halo is calculated using methods described in Section \ref{grav_lensing} (see Wright and Brainerd \cite{lensing_calc} for far more detail). The total convergence and shear along the LOS of the lightcone is simply the sum of the convergence and shear contributions of each halo contained in the lightcone, as well as the additional convergence terms from the smooth-component correction. The \texttt{Pangloss} predicted lensed source ellipticity is then calculated using Equation \eqref{lensing}.

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Pangloss Lightcone Data Plots
\begin{figure*}
    \centering
    \includegraphics[width=0.75\textwidth]{figs-swe/thesis/lightcone_plots.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:lightcone}
\end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

Figure \ref{fig:lightcone} displays different views an example lightcone where the foreground halos are plotted as circles whose size are proportional to their galaxy brightness (top left), angular size (top center), convergence contributions (top right), and halo mass (bottom). Figure \ref{fig:subplots} shows a few examples of mass structures in the foreground catalog along with the ray-traced convergence (plotted in inverse greyscale) and shear maps (plotted as red ellipticity sticks) from \cite{ray_tracing}.

Instead of comparing the ray-traced and \texttt{Pangloss} predicted ellipticities for individual galaxies, the lensing is characterized globally with correlation functions. As described in section \ref{corr_functions}, the ellipticity-ellipticity correlation function measures how correlated the ellipticities of pairs of galaxies are as a function of separation distance, while the galaxy-mass correlation function measures the correlation of lensed ellipticities around foreground halos as a function of separation distance. Both correlation functions are used in this work to estimate how well the \texttt{Pangloss} framework models weak lensing by dark matter structures and are computed using the publicly available \texttt{TreeCorr} module written by Mike Jarvis\footnote{https://github.com/rmjarvis/TreeCorr}. Note that the ellipticity-ellipticity correlation function definition used in \texttt{TreeCorr} is slightly different than that used in most of the literature; for a derivation of the connection between Jarvis's definition and the more common Schneider definition \cite{schneider}, see Appendix (??).

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Two or Three subplots of foreground/background/shear/kappa maps
% \begin{figure*}
%     \centering
%     % Volume
%     \begin{subfigure}[H]{0.415\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/thesis/vol_densities.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{CAPTION}
%         \label{fig:volume_densities}
%     \end{subfigure}
%     ~
%     % Surface
%     \begin{subfigure}[H]{0.425\textwidth}
%         \includegraphics[width=\textwidth]{figs-swe/thesis/surf_densities.png}
%         \captionsetup{justification=raggedright,singlelinecheck=false}
%         \caption{CAPTION}
%         \label{fig:surface_densities}
%     \end{subfigure}
%     \caption{}
%     \label{fig:subplots}
% \end{figure*}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
% Ellipticity-Ellipticity Correlation
\begin{figure}
    \centering
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/gg_corr.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        %The ellipticity-ellipticity correlation function for ?? sources at separation distances from 0.1 to 2 arcminutes and a lightcone radius of $R=2$ arcminutes. While the correlation of the intrinsic ellipticities before lensing (blue) shows no correlation, the correlation of the ray-traced ellipticities (green) and halo model predicted ellipticities (purple) both show significant positive correlation on scales up to 1 arcminute. The halo model predicts a small negative correlation above this scale, while the ray-tracing shows positive correlation on all scales.}
        %The correlation of the intrinsic ellipticities before lensing is plotted in blue, and shows no correlation. The correlation of the true ray-traced ellipticities is plotted in green and shows significant positive correlation on all scales. The halo predicted correlation, plotted in purple, shows significant positive correlation until a separation distance of 0.8 arcminutes.
        \label{gg_corr}
    \end{subfigure}
    ~~
    \begin{subfigure}{0.475\textwidth}
        \includegraphics[width=\textwidth]{figs-swe/gg_corr_series.png}
        \captionsetup{justification=raggedright,singlelinecheck=false}
        \caption{CAPTION}
        %The ellipticity-ellipticity correlation functions for a series of halo model frameworks with increasing lightcone radii from $R=1$ to $R=8$ arcminutes, all plotted in purple with various line styles. These are compared to the same ray-traced correlation from Figure \ref{gg_corr}.}
        \label{gg_corr_series}
    \end{subfigure}
    \caption{}
\end{figure}
%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% Galaxy-mass correlation function
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/ng_corr.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    %The galaxy-mass correlation function of 3,240 sources at separation distances from 0.1 to 2 arcminutes and a lightcone radius of $R=2$ arcminutes. There is no significant correlation of the intrinsic ellipticities (blue) as expected, and both the ray-traced correlation (green) and halo model correlation (purple) are positive on all scales. The green and purple curves are in agreement from 0.2 to 0.7 arcminutes, but have significant discrepancies on scales above and below this range.}
    \label{ng_corr}
\end{figure}
%
% ...Same progression corr function as gg??
%
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
%% 3 panel plot of binned Hilbert map, Pangloss map, and difference
\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figs-swe/thesis/kappamaps.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{CAPTION}
    \label{fig:kappamaps}
\end{figure*}
%
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

%%------------------------------------------------------------------------------------------------------------------------------------------
% Mean lightcone CPU time
\begin{figure}
    \centering
    \includegraphics[width=0.475\textwidth]{figs-swe/thesis/profile_bar.png}
    \captionsetup{justification=raggedright,singlelinecheck=false}
    \caption{OLD: The mean CPU runtime per lightcone for various runs with different (and cumulative) code optimizations described in the text. The cumulative changes account for a ??\% decrease in the mean CPU time and a ?? of the variation. Currently, the prediction can be made for a single lightcone in\\$??\pm??$ ms.}
    \label{fig:cpu_plot}
\end{figure}
%%------------------------------------------------------------------------------------------------------------------------------------------

@@

While the current \texttt{Pangloss} framework can handle the lensing predictions of ?? sources in a ?? arcmin$^2$ field comfortably on a single processor, the goal is to scale up the framework to make predictions for background sources of the same number density in a 100 deg$^2$ field; this will require 3.6 million lightcones, each containing at least ??? relevant foreground halos. Luckily, the prediction is trivially parallelizable as the shear and convergence calculation of each lightcone is independent of all other lightcones. This makes GPUs an attractive candidate for future work, as it would only take 360 GPUs with 10,000 threads each to carry out the prediction. Additionally, the ?? MB of RAM per lightcone is sufficiently small to fit 10,000 lightcones on a GPU.\\

(More??)

@@

% OLD: I would like to thank the Department of Energy for funding this research and giving me this wonderful opportunity, as well as SLAC National Laboratory for hosting me.

%This work was greatly aided by numerous conversations with Risa Wechsler from Stanford University and Matt Becker from the Kavli Institute for Particle Astrophysics and Cosmology (KIPAC), as well as Jesus Pando from DePaul University. Finally, I am incredibly grateful to Phil Marshall for his unending support, patience, and guidance throughout this research.
% This work was supported in part by the U.S. Department of Energy, Office of Science, Office of Workforce Development for Teachers and Scientists (WDTS) under the Science Undergraduate Laboratory Internships Program (SULI).

@@

The definition for the shear-shear (or equivalently, in the case of weak lensing, ellipticity-ellipticity) correlation function components in Mike Jarvis's \texttt{TreeCorr} package is different from the common definition given in Equation \eqref{gg_corr_def} from Schneider \cite{schneider}. For the aid of those who wish to use this very useful package but still use the conventional correlation function defintion, I have outlined the connection between the two definitions and the outputs of \texttt{TreeCorr} below:
